---
title: "Homework 05"
author: "Adam Guerra"
format: html
editor: visual
date: 05/30/2023
html: 
  toc: TRUE
  toc-location: left
  code-fold: true
  theme: sandstone
execute: 
  message: false
  warning: false
---

```{r, include = FALSE, messgae = FALSE}
#load in packages
library(tidyverse)
library(car)
library(sqldf)
library(janitor)
library(ggthemes)
library(naniar)
library(performance)
library(broom)
library(ggeffects)
library(flextable)
library(naniar)
library(corrplot)
library(AICcmodavg)
library(GGally)
library(here)

#load in data
plant <- read_csv(here("data/sarracenia.csv")) |> 
        clean_names()

#subset desired columns
plant_clean <- plant |> 
               select(totmass, species, feedlevel, sla, 
                      chlorophyll, amass, num_lvs, num_phylls)
```

## Introduction

Sarracenia is a genus of carnivorous plants typically found on the eastern seaboard of the United States. Within our data there are ten different species.

## Methods

#### Protocols

"Two plants of each Sarracenia species were assigned to one of six feeding levels in a regression design ranging from 0 - 0.25g of finely ground wasps per feeding (for small species), 0 - 0.5g (for species of intermediate sizes), and 0 - 1.0g (for large species); N equals120 plants total. Aboveground size and Amass were measured prior to commencing treatments. Plants were fed once/week for 7 weeks." (Ellison).

#### Organization

The data is organized into columns that are characteristics, with each row being a an individual plant with its corresponding observations. There are 120 records and 32 variables in CSV format. 11 of the columns were calculated, while the other 21 were measurements.

#### Missing Data

```{r}
#visualize missing observations
gg_miss_var(plant_clean) +
  labs(title = "Missing Values in the Dataset", 
       caption = "Figure 1: There are a high number of mising values in the chlorophyll content and mass-based light-saturated 
       photosyntehtic rate of youngest leaf columns. There are also quite a few values missing from the specific 
       leaf area of the youngest fulley expanded leaf column.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        plot.caption.position = "plot")
```

If chlorophyll content is missing, the amass column likely cannot be calculated. This could be the reason that each of those columns have the same number of missing N/A. Although it could also be that those specific plants did not meet the requirements to take the measurement in other categories as well within the full data set.

#### Subset NAs

```{r}
#drop NAs
plant_sub <- plant_clean |> 
             drop_na(sla, chlorophyll, num_phylls, amass, num_lvs)
```

#### Correlation

```{r}
#calculate phearsons
plant_corr <- plant_sub |> 
              select(1,3:8) |> 
              cor(method = "pearson")

#visualize pearsons correlation 
corrplot(plant_corr, method = "color", addCoef.col = "black")
```

## Results
